The Effect of Events on Australian Federal Parliamentary Discussion (1901--2017)

### 2018-10-14 outputs/docs/events/paper.Rmd
UPdate paper

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Events on Australian Parliamentary Discussion (1851--2017)"
thanks: "Thank you to John Tang, Zach Ward, Tim Hatton, Martine Mariotti, Tianyi Wang, Matt Jacob, Leslie Root, Jill Sheppard, Matthew Kerby, Chris Cochrane and Dan Simpson for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how parliamentary discussion changes in response to different types of events in Australian history. We first create a dataset of what was said in Australian state and federal parliaments from the mid-1800s through to 2017 based on available public records. To reduce the dimensionality of this dataset we use a topic model that allows for correlation and steady changes over time. We analyse the effect of various events using a Bayesian hierarchical Dirichlet model. We find that: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. Our findings have implications for how we think about the longer-term trajectory of government policymaking as the media cycle becomes increasingly focused on short-term events."
keywords: "text analysis, politics, Australia"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no
---

@@

yearsUsed <- read_csv(here::here("outputs/misc/years_used.csv"), col_types = cols()) %>% 
  replace_na(list(Parliament = ""))
elections <- read_csv(here::here("inputs/misc/misc_elections_data.csv"), col_types = cols())
governments <- read_csv(here::here("inputs/misc/change_of_pm.csv"), col_types = cols())
keyEvents <- read_csv(here::here("inputs/misc/key_events.csv"), col_types = cols())
significances <- read_csv(here::here("outputs/misc/significance.csv"), col_types = cols())
top_words <- read_csv2(here::here("outputs/misc/top_words_40.csv"), col_types = cols())
missing_days <- read_csv(here::here("outputs/misc/missing_days.csv"), col_types = cols())
hansard_sources <- read_csv(here::here("outputs/misc/hansard_sources.csv"), col_types = cols())

@@

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events--both expected, such as an election, and unexpected, such as a recession or terrorist attack--can affect the course of parliamentary discussion. For instance, think of how a new government often goes to some trouble to appear different to the one they replace, or how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in Australian state and federal parliaments. Our earliest record is from 1851 and we consider the time period through to 2017. 
<!-- The dataset is detailed in Section \@ref(data).  -->
We use a topic model for dimensionality reduction and to allow for correlation between topics and steady changes over time. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government or elections; changes in economic conditions; and other significant events such as the 9/11 attacks. 
<!-- The models are detailed in Section \@ref(model). -->

<!-- Section \@ref(results) details our results, but broadly  -->
We find: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. **[More here on the nature of the differences.]**

In this paper we apply a topic model to a dataset of larger-scale parliamentary text records from multiple Australian parliaments and then use a Bayesian hierarchical Dirichlet model to examine events. From a data perspective we contribute **[TBD]**. From a methods perspective, we contribute **[TBD]**. Our model can be used in many settings where the results of topic models need to be further analysed.

Our work is cross-disciplinary and in the traditions of political economy, economic history, political science and statistics. It contributes to a growing modern quantitative social sciences literature that considers text as an input to more traditional methods, rather than requiring separate analysis. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in this approach to only increase.

@@

# Data
## Parliamentary text data
Following the example of the UK a daily text record called Hansard of what was said in Australian parliaments has been made available since their establishment.[^Verbatimfootnote] Earlier work on the influence of parliaments, such as @ZandenBuringhBosker2012, often examined broader activity measures such as counts of sitting days. This allowed for long time frames and wide comparisons. But analysing Hansard records and their equivalents directly are an increasingly viable and popular source of data as new methods and reduced computational costs make larger-scale analysis easier, complementing other measures.

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, allowed @RheaultCochran2018 to examine ideology and party polarisation in Britain and Canada. In the UK, @Duthie2016 analysed Hansard records to examine which politicians made supportive or aggressive statements toward other politicians between 1979 and 1990 and @PetersonSpirling2018 examined polarisation. And as digitisation methods improve older UK records can be analysed, for instance @Dimitruk2018 considers the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2017 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US @GentzkowShapiroTaddy2018 examine congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

[^Verbatimfootnote]: While Hansard is not necessarily verbatim, it is considered close enough for text-as-data purposes. For instance, @Mollin2008 found that in the case of the UK Hansard the differences would only affect specialised linguistic analysis. @Edwards2016 examined Australia, New Zealand and the UK, and found that changes were usually made by those responsible for creating the Hansard record, instead of the parliamentarians.

Australian Hansard records have been analysed for various purposes. For instance, @Rasiah2010 examined Hansard records for the Australian House of Representatives to examine whether politicians attempted to evade questions about Iraq during February and March 2003. And @GansLeigh2012 examined Australian Hansard records to associate mentions by politicians of certain public intellectuals with neutral or positive sentiment.

Australian parliaments generally make their daily Hansard records available online as PDFs and these are considered the official release. Additionally, XML records are available in some cases.[^XMLfootnote] We detail the sources of our Hansard PDFs in \@ref(hansardsources) and Appendix \@ref(examplehansardpage) provides an example of what a Hansard PDF looks like. There are roughly 54,331 **[UPDATE]** days worth of publicly available Hansard records across the state and federal parliaments (Table \ref{tab:yearsusedtable}). Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

@@

```{r yearsusedtable, echo = FALSE, results = 'asis'}

yearsUsed %>%
  kable(
    booktabs = T,
    align = c('l', 'r', 'r', 'r'),
    caption = "Hansard records used") %>%
  kable_styling(font_size = 12)

The formatting of the Hansard records changes between the different parliaments and over time. We use scripts written in R (@R2018) to convert the PDFs into daily text records. An example of the workflow and some reduced-detail scripts are provided in Appendix \@ref(hansardworkflow). Some error is introduced at this stage because many of the records are in a two-column format that need to be separated, and the PDF parsing is not always accurate especially for older records. An example of the latter issue is that 'the' is often parsed as 'thc'. These errors are corrected when they occur at scale and can be identified.

The percentage of stop-words in each record is reasonably consistent over time. This suggests that there is no significant difference in the quality of the parsing over time. Details of this process are provided in Appendix \@ref(stopwordsgraph). We mainly use Hansard records on a daily basis in this paper. Text is usually pre-processed before topic models are used. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multiword names titles and phrases, such as new zealand to new_zealand. Then the sentences are deconstructed and each word considered individually.

@@

The goal of our modeling strategy is twofold. Firstly, we want to topic modeling __(REF)__ to summarise all text contained in Hansard into meaningful topics, which reduce the dimensionality of the text data and capture the main themes discussed in parliament over time. Secondly, we want to relate the resulting topic distributions to temporal trends, changes and events, such as governement changes, elections, and other external events.

We use a Correlated Topic Model (CTM) __(REF)__ to obtain estimated topic distributions over time. We consider these topic distributions as reduced dimension inputs that can be analysed within another model. We then formulate a Bayesian hierarchical Dirichlet model to assess changes in the topic distributions in relation to events of interest. __In Appendix \@ref(word2vec), we include an alternative approach that follows @Taddy2015 by using word2vec, which more closely uses words, rather than topics, as an input.__

@@

## Correlated Topic Model
As mentioned in the previous section, one of the limitations of LDA is that the model assumes that the presence of one topic is not correlated with the presence of another topic. Indeed, the corrleation between topics is not modeled or accounted for. However, in relaity, we would expect Hansard topics related the army may be more commonly found with topics related to the navy, but less commonly with topics related to banking. The goal of the CTM __(REF)__ is to account for this correlation between topics, in order to produce more realistic and stable topic distributions over time.

@@

However, the replacement of the Dirichlet distribution, which is conjugate to the multinomial, with the logistic Normal distribution adds a level of computational complexity to CTM. The posterior distributions of the parameters of interest ($\beta_{1:K}, \theta_{1:D}, z_{1:D, 1:N}$) can no longer be obtained using standard simulation techniques such as Gibbs Sampling. __(REF)__ develop a fast variational inference procedure for estimating the CTM. This estimation approach has also been extended by @RobertsStewartAiroldi2016 as part of their work on Structural Topic Models.

@@

The main output of the CTM of interest is the proportion of each topic appearing in each document, that is, the $\theta_d$. For the next stage of the modeling process, we are interested in how the distribution of topics changes in relation to different types of events. But with around 40 topics **[UPDATE]** for each of the roughly 54,331 **[UPDATE]** sitting days across the parliaments and houses the data are still too noisy to easily visualise changes around events.

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Events on Australian Federal Parliamentary Discussion (1901--2017)"
thanks: "Thank you to John Tang, Zach Ward, Tim Hatton, Martine Mariotti, Tianyi Wang, Matt Jacob, Leslie Root, Jill Sheppard, Matthew Kerby, Chris Cochrane and Dan Simpson for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "We systematically analyse how parliamentary discussion changes in response to different types of events in Australian history. We first create a dataset of what was said in the Australian Federal parliament from 1901 through to 2017 based on available public records. To reduce the dimensionality of this dataset we use a correlated topic model. We analyse the effect of various events using a Bayesian hierarchical Dirichlet model. We find that: 1) changes in government tend be associated with topic changes even when the party in power does not change; 1) elections that do not result in a change in government are rarely associated with topic changes; 2)  3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. Our findings have implications for how we think about the longer-term trajectory of government policymaking as the media and political cycle becomes increasingly focused on short-term events."
keywords: "text analysis, politics, Australia"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no
---

@@

elections <- read_csv(here::here("inputs/misc/misc_elections_data.csv"), col_types = cols())
governments <- read_csv(here::here("inputs/misc/change_of_pm.csv"), col_types = cols())
keyEvents <- read_csv(here::here("inputs/misc/key_events.csv"), col_types = cols())
significances <- read_csv(here::here("outputs/misc/significance.csv"), col_types = cols())
top_words <- read_csv2(here::here("outputs/misc/top_words_40.csv"), col_types = cols())
missing_days <- read_csv(here::here("outputs/misc/missing_days.csv"), col_types = cols())
hansard_sources <- read_csv(here::here("outputs/misc/hansard_sources.csv"), col_types = cols())

@@

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events--both expected, such as an election, and unexpected, such as a recession or terrorist attack--can affect the course of parliamentary discussion. For instance, think of how a new government often goes to some trouble to seem different to the one they replace, or how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in Australian's federal parliament. We create a dataset based on all sitting days from 1901 through to 2017. Over this period we have records for 7,869 days in the House of Representatives and 6,682 days in the Senate, although we drop a handful of days for various reasons. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lowercase, remove punctuation and numbers and join commonly collocated words.

We use a topic model that allows correlation between topics to reduce the dimensionality of our dataset. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government; elections; changes in economic conditions; and other significant events such as the 9/11 attacks.

The advantage of using an analysis model in this way is that it: allows the events to have more-complicated autocorrelated functional forms; allows us to implement pooling across groups of similar documents; and it identifies outlying topic distributions without the need to pre-specify the event of interest. The main weakness of this approach is that it does not appropriately propagate uncertainty and future models should improve on this.

We find: 1) changes in government tend be associated with topic changes even when the party in power does not change. For instance, the change from Menzies to Holt in 1966; the change from Hawke to Keating in 1991; and the change from Rudd to Gillard in 2010 are all associated with significant changes in topics despite no change in the party in power. 2) As expected, elections where the party in power also changes, such as Whitlam in 1972, Hawke in 1983, Howard in 1996, Rudd in 2007, and Abbott in 2013 are associated with topic changes, but the 2001 and 2004 Howard re-elections stand out as elections where the government was returned yet there was a significant change in topics. 3) Economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. **[More here on that one.]** **[Do we have different slopes that we could talk about?]**

Our work contributes to a growing modern quantitative social sciences literature that considers text as an input to more traditional methods, rather than requiring separate analysis. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in this approach to only increase.

@@

# Data
## Parliamentary text data
Following the example of the UK a daily text record called Hansard of what was said in Australian parliaments has been made available since their establishment.[^Verbatimfootnote] Earlier work on the influence of parliaments, such as @ZandenBuringhBosker2012, often examined broader activity measures such as counts of sitting days. This allowed for long time-frames and wide comparisons. But analysing Hansard records and their equivalents directly is increasingly viable as new methods and reduced computational costs make it easier.

[^Verbatimfootnote]: While Hansard is not necessarily verbatim, it is considered close enough for text-as-data purposes. For instance, @Mollin2008 found that in the case of the UK Hansard the differences would only affect specialised linguistic analysis. @Edwards2016 examined Australia, New Zealand and the UK, and found that changes were usually made by those responsible for creating the Hansard record, instead of the parliamentarians. As those who create Hansard are tasked with creating an accurate record of proceedings, this suggests the records should be fit for the purpose of our analysis.

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, allowed @RheaultCochran2018 to examine ideology and party polarisation in Britain and Canada. In the UK, @Duthie2016 analysed Hansard records to examine which politicians made supportive or aggressive statements toward other politicians between 1979 and 1990 and @PetersonSpirling2018 examined polarisation. And as digitisation methods improve older UK records can be analysed, for instance @Dimitruk2018 considers the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2017 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US @GentzkowShapiroTaddy2018 examined congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

Parts of Australian Hansard records have been analysed for various purposes. For instance, @Rasiah2010 examined Hansard records for the Australian House of Representatives to examine whether politicians attempted to evade questions about Iraq during February and March 2003. And @GansLeigh2012 examined Australian Hansard records to associate mentions by politicians of certain public intellectuals with neutral or positive sentiment.

Australian parliaments generally make their daily Hansard records available online as PDFs and these are considered the official release. Additionally, XML records are available in some cases.[^XMLfootnote] We detail the sources of our Hansard PDFs in \@ref(hansardsources) and Appendix \@ref(examplehansardpage) provides an example of what a Hansard PDF looks like. There are 14,551 days worth of publicly available Hansard records across the houses of federal parliament. Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

@@

The formatting of the Hansard records changes over time. We use scripts written in R (@R2018) to convert the PDFs into daily text records. An example of the workflow and some reduced-detail scripts are provided in Appendix \@ref(hansardworkflow). Some error is introduced at this stage because many of the records are in a two-column format that need to be separated, and the PDF parsing is not always accurate especially for older records. An example of the latter issue is that 'the' is often parsed as 'thc'. These errors are corrected when they occur more than twice and can be identified.

The percentage of stop-words in each record is reasonably consistent over time. This suggests that there is no significant difference in the quality of the parsing over time. Details of this check are provided in Appendix \@ref(stopwordsgraph). We use Hansard records on a daily basis in this paper. We pre-process our text before applying a topic model. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multiword names titles and phrases, such as new zealand to new_zealand. Then the sentences are deconstructed and each word considered individually.

@@

The goal of our modeling strategy is twofold. Firstly, we want to use topic modeling [@Blei2003latent] to summarise the Hansard text into meaningful topics, which reduce the dimensionality of the text data and capture the main themes discussed in parliament over time. Secondly, we want to relate the resulting topic distributions to temporal trends, changes and events, such as governement changes, elections, and other external events.

We use a Correlated Topic Model (CTM) @BleiLafferty2007 to obtain estimated topic distributions over time. We consider these topic distributions as reduced dimension inputs that can be analysed within another model. We then formulate a Bayesian hierarchical Dirichlet model to assess changes in the topic distributions in relation to events of interest. In Appendix \@ref(word2vec), we include an alternative approach that follows @Taddy2015 by using word2vec, which more closely uses words, rather than topics, as an input.

@@

## Correlated Topic Model
As mentioned in the previous section, one of the limitations of LDA is that the model assumes that the presence of one topic is not correlated with the presence of another topic. Indeed, the corrleation between topics is not modeled or accounted for. However, in relaity, we would expect Hansard topics related the army may be more commonly found with topics related to the navy, but less commonly with topics related to banking. The goal of the CTM [@BleiLafferty2007] is to account for this correlation between topics, in order to produce more realistic and stable topic distributions over time.

@@

However, the replacement of the Dirichlet distribution, which is conjugate to the multinomial, with the logistic Normal distribution adds a level of computational complexity to CTM. The posterior distributions of the parameters of interest ($\beta_{1:K}, \theta_{1:D}, z_{1:D, 1:N}$) can no longer be obtained using standard simulation techniques such as Gibbs Sampling. @BleiLafferty2007 develop a fast variational inference procedure for estimating the CTM. This estimation approach has also been extended by @RobertsStewartAiroldi2016 as part of their work on Structural Topic Models.

@@

The main output of the CTM of interest is the proportion of each topic appearing in each document, that is, the $\theta_d$. For the next stage of the modeling process, we are interested in how the distribution of topics changes in relation to different types of events. But with 40 topics for each of the roughly 14,551 **[UPDATE]** house-sitting-days the data are still too noisy to easily visualise changes around events.
